---
title: "Example Analysis"
code-fold: true
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

In this vignette, I will analyze a dataset about estimated salary for college students. The goal here is to demonstrate the functions from the R package `bbplot` and wrangle and plot the example data from TidyTuesday.

* Question: How does the early career pay differ across the U.S.?

* Audience: R package `bbplot` users who are interested in seeing example data wrangling and visualizations.

* Dataset: I obtained the original dataset from  [TidyTuesday](https://github.com/rfordatascience/tidytuesday/tree/master/data/2020/2020-03-10).

* Data dictionary: I referred to this [data dictionary](https://github.com/rfordatascience/tidytuesday/tree/master/data/2020/2020-03-10#data-dictionary) from TidyTuesday.

## Data Analysis

#### Step 1: load packages 
```{r bg, message = FALSE, warning = FALSE}
library(bbplot)
library(here)
library(readr)
library(tidyverse)
library(ggplot2)
library(maps)
library(purrr)
```

#### Step 2: load data
```{r load, message = FALSE, warning = FALSE}
# check if a directory named data exists locally
if (!dir.exists(here("data"))) {
  # if not, create it programmatically
  dir.create(here("data"))
}

# check whether we have the input data
if (!file.exists(here("data", "salary_potential.rda"))) {
    # If we don't, then download the data
    salary_potential <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/salary_potential.csv")

    save(salary_potential, file = here("data", "salary_potential.rda"))
}

# load the data
load(here("data", "salary_potential.rda"))
```

#### Step 3: data wrangling
```{r}
# check the content of the data set
head(salary_potential)
map_chr(salary_potential, .f = typeof)
map_int(salary_potential, .f = ~ length(unique(.x)))
```

```{r}
# calculate the average estimated early career pay
salary_st_avg <- salary_potential %>%
  select(name, state_name, early_career_pay) %>%
  rename(region = state_name) %>%
  mutate(region = tolower(region)) %>%
  group_by(region) %>%
  summarise(
    avg_early_career_pay = mean(early_career_pay)
  )
  
# find coordinates for us states
us_map <- map_data("state") %>%
  mutate(region = gsub(" ", "-", region))

# add salary info to the map
map_salary_st_avg <- us_map %>%
  left_join(salary_st_avg, by = join_by(region))
```

```{r}
# select salary info from states of interest
salary <- salary_potential %>%
  select(state_name, early_career_pay, mid_career_pay) %>%
  rename(state = state_name) %>%
  mutate(state = tolower(state)) 

# break the data set down into groups defined by states
by_state <- split(salary, salary$state)

# calculate average early pay and mid pay for each state
early_pay <- map_dbl(by_state, ~ mean(.x$early_career_pay, na.rm=TRUE))
mid_pay <- map_dbl(by_state, ~ mean(.x$mid_career_pay, na.rm=TRUE))

# combine average pay into a table
pay_trend <- tibble(
  state = names(early_pay),
  avg_early_pay = early_pay,
  avg_mid_pay = mid_pay
)
```

```{r}
# select salary info from states of interest
pay_stem_trend <- salary_potential %>%
  rename(state = state_name) %>%
  mutate(state = tolower(state)) %>%
  filter(state %in% c("california", "michigan", "maryland", "washington"))
```

#### Step 4: data visualization
```{r fig.dim = c(8,6)}
# plot the average estimated early career pay across the US
map_salary_st_avg %>% ggplot() +
  geom_polygon(aes(x = long, y = lat, group = group, fill = avg_early_career_pay)) +
  scale_fill_gradient(low = "#FAAB18", high = "#1380A1") +  # Adjust the color scale
  labs(title = "The Average Estimated Early Career Salary Across the U.S.", subtitle = str_wrap("California, Texas, and the states in the northeastern region have higher estimated early career pay compared to states in the central portion of the U.S.", 60), caption = "Yuhan Xiao", fill = "Salary") +
  theme_void(base_size = 14) +
  theme(plot.title = element_text(hjust = 0.5, vjust = 1, size = 15), 
        plot.subtitle = element_text(hjust = 0.5),
        plot.background = element_rect(fill = "white"),
        text = element_text(family = "Times"))
```

```{r fig.dim = c(8,6)}
# plot the association between average early career pay to average mid career pay
pay_trend_plot <- pay_trend %>% 
  ggplot() +
  geom_point(aes(x = avg_early_pay, y = avg_mid_pay)) +
  labs(title = str_wrap("The Association Between Average Estimated Early Career Salary and Mid Career Salary by State", 30), subtitle = str_wrap("A higher early career salary is associated with a higher mid career salary", 40), x = "average early career salary", y = "average mid career salary", caption = "Yuhan Xiao") +
  bbc_style() 

finalise_plot(plot_name = pay_trend_plot,
  source = "Source: TidyTuesday",
  save_filepath = here("data","pay_trend-nc.png"),
  width_pixels = 640,
  height_pixels = 550,
  logo_image_path = here("data","placeholder.png"))
```

::: {.callout-note}
* Note: I have added captions and axis labels. This plot, however, does not show these components because it follows the BBC style.
:::

```{r warning = F, message = F, fig.dim = c(8,6)}
pay_stem_trend %>% ggplot(aes(x=stem_percent, y = early_career_pay, color = state)) +
  geom_point() +
  geom_smooth() + 
  facet_wrap(~state, nrow = 2) +
  labs(title = str_wrap("The Association Between Average Estimated Early Career Salary and Percent of STEM students by State", 60), subtitle = str_wrap("A higher percent of student body in STEM is associated with a higher estimated early career salary", 60), x = "STEM percent", y = "average early career salary", caption = "Yuhan Xiao") +
  theme_minimal(base_size = 14) +
  theme(plot.title = element_text(hjust = 0.5, vjust = 1, size = 15), 
        plot.subtitle = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 0, hjust = 1),
        plot.background = element_rect(fill = "white"),
        text = element_text(family = "Times"),
        legend.position = "none")
```

## Summary

Colleges have higher estimated early career salary for students if they are located in California, Texas, the northeastern region or if they have a higher percent of student body in STEM. A higher early career salary is associated with a higher mid career salary.

## Functions
-   `bbplot`: `bbc_style()`, `finalise_plot()`
-   `dplyr`: `select()`, `group_by()`, `summarise()`, `mutate()`, `rename()`, `filter()`
-   `stringr`: `str_wrap()`
-   `purrr`: `map_dbl()`, `map_chr()`, `map_int()`
-   `ggplot2`: `geom_polygon()`, `geom_point()`, `geom_smooth()`, `facet_wrap()`
