---
title: "Example Analysis"
code-fold: true
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

In this vignette, I will analyze a dataset about estimated salary for college students. The goal here is to demonstrate the functions from the R package `bbplot` and wrangle and plot the example data from TidyTuesday.

* Question: How does the early career pay differ across the U.S.?

* Audience: R package `bbplot` users who are interested in seeing example data wrangling and visualizations.

* Dataset: I obtained the original dataset from  [TidyTuesday](https://github.com/rfordatascience/tidytuesday/tree/master/data/2020/2020-03-10).

* Data dictionary: I referred to this [data dictionary](https://github.com/rfordatascience/tidytuesday/tree/master/data/2020/2020-03-10#data-dictionary) from TidyTuesday.

## Data Analysis

#### Step 1: load packages 
```{r bg, message = FALSE, warning = FALSE}
library(bbplot)
library(here)
library(readr)
library(tidyverse)
library(ggplot2)
library(maps)
library(purrr)
```

#### Step 2: load data
```{r load, message = FALSE, warning = FALSE}
# check if a directory named data exists locally
if (!dir.exists(here("data"))) {
  # if not, create it programmatically
  dir.create(here("data"))
}

# check whether we have the input data
if (!file.exists(here("data", "salary_potential.rda"))) {
    # If we don't, then download the data
    salary_potential <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/salary_potential.csv")

    save(salary_potential, file = here("data", "salary_potential.rda"))
}

# load the data
load(here("data", "salary_potential.rda"))
```

#### Step 3: data wrangling
```{r}
# calculate the average estimated early career pay
salary_st_avg <- salary_potential %>%
  select(name, state_name, early_career_pay) %>%
  group_by(state_name) %>%
  summarise(
    avg_early_career_pay = mean(early_career_pay)
  ) %>%
  rename(region = state_name) %>%
  mutate(region = tolower(region))

# find coordinates for us states
us_map <- map_data("state") %>%
  mutate(region = gsub(" ", "-", region))

# add salary info to the map
map_salary_st_avg <- us_map %>%
  left_join(salary_st_avg, by = join_by(region))
```

```{r}
```

#### Step 4: data visualization
```{r fig.dim = c(8,6)}
# plot the average estimated early career pay across the US
map_salary_st_avg %>% ggplot() +
  geom_polygon(aes(x = long, y = lat, group = group, fill = avg_early_career_pay)) +
  scale_fill_gradient(low = "#FAAB18", high = "#1380A1") +  # Adjust the color scale
  labs(title = "The Average Estimated Early Career Salary Across the U.S.", subtitle = str_wrap("California, Texas, and the states in the northeastern region have higher estimated early career pay compared to states in the central portion of the U.S.", 60), caption = "Yuhan Xiao", fill = "Salary") +
  theme_void(base_size = 14) +
  theme(plot.title = element_text(hjust = 0.5, vjust = 1, size = 15), 
        plot.subtitle = element_text(hjust = 0.5),
        plot.background = element_rect(fill = "white"),
        text = element_text(family = "Times"))
```

## Summary

## Functions
-   `bbplot`: `bbc_style()`, `finalise_plot()`
-   `dplyr`: `select()`, `group_by()`, `summarise()`, `mutate()`, `rename()`
-   `stringr`: `str_wrap()`
-   `purrr`: 
-   `ggplot2`: `geom_polygon()`, `facet_wrap()`
